Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-08-13 Report/conclusions.tex
modified for review

In this work, we presented the \acl{QM} objective to minimize the number of backlogged packets by designing the transmit precoders across the space-frequency dimension in a multi-cell \ac{MU} \ac{MIMO} system. We proposed the \ac{JSFRA} scheme, which adopts the \ac{SCA} technique to model the nonconvex constraint as a convex constraint in an iterative manner to design the precoders for the \acl{QM} objective. We also proposed an alternative approach using the \ac{MSE} relaxation for the same problem when the receive beamformers are based on the \ac{MMSE} receivers. The exponent used in the proposed formulation can be modified to achieve either greedy allocation or equally fair allocation. We also proposed the distributed solutions for the proposed centralized \ac{JSFRA} problem using the \ac{ADMM} method, which has better convergence behaviour compared to the primal decomposition. In addition, we also proposed an iterative algorithm based on the \ac{KKT} conditions for the \ac{MSE} reformulated \ac{JSFRA} scheme to design the precoders in a distributed manner using the closed form solutions.

>>>>

In this work, we presented the \acl{QM} objective to minimize the number of backlogged packets by designing the transmit precoders across the space-frequency dimension in a multi-cell \ac{MU} \ac{MIMO} system. We proposed the \ac{JSFRA} scheme, which adopts the \ac{SCA} technique to model the nonconvex constraint as a convex constraint in an iterative manner to design the precoders for the \acl{QM} objective. We also proposed an alternative approach using the \ac{MSE} relaxation for the same problem when the receive beamformers are based on the \ac{MMSE} receivers. The exponent used in the proposed formulation can be modified to achieve either greedy allocation or equally fair allocation. We also proposed the distributed solutions for the proposed centralized \ac{JSFRA} problem using the \ac{ADMM} method, which has better convergence behavior compared to the primal decomposition. In addition, we also proposed an iterative algorithm based on the \ac{KKT} conditions for the \ac{MSE} reformulated \ac{JSFRA} scheme to design the precoders in a distributed manner using the closed form solutions.

### 2014-08-13 Report/decomposition_solutions.tex
updated kkt,wsrm,admm,primal
Corrections after first reading
Signed-off-by: Ganesh <gvenkatr@ee.oulu.fi>
modified for review

This section addresses the distributed precoder designs for the proposed \ac{JSFRA} scheme. The formulation in \eqref{eqn-9} or \eqref{eqn-mse-2} requires a centralized controller to perform the precoder design for all users belonging to the coordinating \acp{BS}. In order to design the precoders independently at each \ac{BS} with the minimal information exchange via backhaul, iterative decentralization methods are considered. In particular, the \ac{PD} and the \ac{ADMM} based \ac{DD} approaches are addressed.

To begin with, let \me{\bar{\mc{B}}_b} denote the set \me{\mc{B} \backslash \{b\}} and \me{\bar{\mc{U}}_b} represents the set \me{\mc{U} \backslash \mc{U}_b}. The centralized problem\footnote{\eqref{eqn-9} for the \ac{SCA} approach and \eqref{eqn-mse-2} for \ac{MSE} reformulation approach} can be equivalently written as
\begin{IEEEeqnarray}{rCl} \label{eqn-decent-1}
\underset{\substack{\gamma_{l,k,n},p_{l,k,n},q_{l,k,n} \\ \mvec{m}{l,k,n}, \beta_{l,k,n}}}{\text{minimize}} &\quad & \sum_{b \in \mc{B}} \| \tilde{\mbf{v}}_b \|_q \IEEEyessubnumber \label{eqn-decent-1a} \\
\text{subject to}&\quad& \eqref{eqn-9.1c} - \eqref{eqn-9.1e} \IEEEyessubnumber,
\end{IEEEeqnarray}
where \me{\tilde{\mbf{v}}_b} denote the vector of of weighted queue deviation corresponding to the users \me{k \in \mc{U}_b}.

Following the similar approach as in \cite{pennanen2011decentralized,tolli2011decentralized}, the coupling constraint\footnote{coupling constraint in \ac{MSE} scheme is given by  \eqref{eqn-mse-2.3}} given by \eqref{eqn-9.1c} can be expressed by grouping the interference contribution from each \acp{BS} in the system as
\begin{IEEEeqnarray}{rCl}\label{eqn-decent-3}
\beta_{l,k,n} & \geq & \sum_{\substack{j = 1\\j \neq l}}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,k,n} |^2 \nonumber \\
&\quad& + \sum_{i \in \mc{U}_{b_k} \backslash \{k\}} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,i,n} |^2 + \sum_{b \in \bar{\mc{B}}_{b_k}} \zeta_{l,k,n,b} \; + \; N_0, \label{eqn-decent-3a}
\end{IEEEeqnarray}
where \me{\zeta_{l,k,n,b}}, which is the total interference caused by the \ac{BS} \me{b} to the \me{\ith{l}} stream of user \me{k \in \mc{U}_{b_k}} on the \me{\ith{n}} sub-channel, is upper bounded by
\begin{equation}
\zeta_{l,k,n,b} \geq \sum_{i \in \mc{U}_b} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2, \forall b \in \bar{\mc{B}}_{b_k}.
\end{equation}

The coupling variable \me{\beta_{l,k,n}} can be decoupled using the variable \me{\zeta_{l,k,n,b}}, which limits the interference caused by the transmission from \ac{BS} \me{b} to the user \me{\ith{k}} corresponding data stream. In order the solve for the global optimal precoders, we need to find the optimal coupling variables \me{\zeta_{l,k,n,b}} by either using \ac{PD} or by using \ac{DD} method. In both approaches, the coupling constraint \eqref{eqn-9.1c} for the \ac{SCA} scheme and \eqref{eqn-mse-2.3} for the \ac{MSE} relaxation scheme are decoupled to perform the distributed precoder design problem.

>>>>

This section addresses the distributed precoder designs for the proposed \ac{JSFRA} scheme. The formulation in \eqref{eqn-9} or \eqref{eqn-mse-2} requires a centralized controller to perform the precoder design for all users belonging to the coordinating \acp{BS}. In order to design the precoders independently at each \ac{BS} with the minimal information exchange via backhaul, iterative decentralization methods are considered. In particular, the \acl{PD} and the \ac{ADMM} based \acl{DD} approaches are addressed.

To begin with, let \me{\bar{\mc{B}}_b} denote the set \me{\mc{B} \backslash \{b\}} and \me{\bar{\mc{U}}_b} represents the set \me{\mc{U} \backslash \mc{U}_b}. In order to study the decomposition based solutions, we consider the solution proposed in the \eqref{eqn-9}, which is based on the Taylor approximation for the nonconvex constraint. The following discussions are equally valid for the \ac{MSE} based solution outlined in \eqref{eqn-mse-2.3} as well. Now, the centralized problem can be equivalently written as
\begin{IEEEeqnarray}{rCl} \label{eqn-decent-1}
\underset{\substack{\gamma_{l,k,n},p_{l,k,n},q_{l,k,n} \\ \mvec{M}{k,n},\mvec{W}{k,n}, \beta_{l,k,n}}}{\text{minimize}} &\quad & \sum_{b \in \mc{B}} \| \tilde{\mbf{v}}_b \|_q \IEEEyessubnumber \label{eqn-decent-1a} \\
\text{subject to}&\quad& \eqref{eqn-9.1c} - \eqref{eqn-9.1e} \IEEEyessubnumber,
\end{IEEEeqnarray}
where \me{\tilde{\mbf{v}}_b} denote the vector of of weighted queue deviation corresponding to the users \me{k \in \mc{U}_b}.

Following the similar approach as in \cite{pennanen2011decentralized,tolli2011decentralized}, the coupling constraint in \eqref{eqn-9.1c} or \eqref{eqn-mse-2.3} can be expressed by grouping the interference contribution from each \acp{BS} in the coordinating set \me{\mathrm{B}} as
\begin{IEEEeqnarray}{rCl}\label{eqn-decent-3}
\beta_{l,k,n} & \geq & \sum_{\substack{j = 1\\j \neq l}}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,k,n} |^2 \nonumber \\
&\quad& + \sum_{i \in \mc{U}_{b_k} \backslash \{k\}} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,i,n} |^2 + \sum_{b \in \bar{\mc{B}}_{b_k}} \zeta_{l,k,n,b} \; + \; N_0\|\mvec{w}{l,k,n}\|^2 \label{eqn-decent-3a}
\end{IEEEeqnarray}
where \me{\zeta_{l,k,n,b}}, which is the total interference caused by the \ac{BS} \me{b} to the \me{\ith{l}} stream of user \me{k \in \mc{U}_{b_k}} on the \me{\ith{n}} sub-channel, is upper bounded by
\begin{equation}
\zeta_{l,k,n,b} \geq \sum_{i \in \mc{U}_b} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2, \forall b \in \bar{\mc{B}}_{b_k}
\end{equation}

The coupling variable \me{\beta_{l,k,n}} can be decoupled using the variable \me{\zeta_{l,k,n,b}}, which limits the interference caused by the transmission from \ac{BS} \me{b} to the user \me{\ith{k}} corresponding data stream. In order the solve for the global optimal precoders, we need to find the optimal coupling variables \me{\zeta_{l,k,n,b}} by either using \acl{PD} or by using \acl{DD} method. In both approaches, the coupling constraint \eqref{eqn-9.1c} for the \ac{SCA} scheme and \eqref{eqn-mse-2.3} for the \ac{MSE} relaxation scheme are decoupled to perform the distributed precoder design problem.

### 2014-08-13 Report/dual_approach.tex
modified for review

In this section, we discuss the \ac{ADMM} decomposition method, which is basically based on the \acf{DD}, but shows better convergence properties. In contrast to the \acl{PD} problem, the \acf{ADMM} method relaxes the interference constraints by including it in the objective function of each subproblem with a penalty pricing \cite{palomar2006tutorial,boyd2011distributed}. In order to decouple the problem \eqref{eqn-decent-1}, the coupling variables \me{\zeta_{l,k,n,b}} in \eqref{eqn-decent-3} are replaced by the respective local copies \me{\mbfa{\zeta}^{\set{b}}, \forall b \in \mc{B}}, which are then solved for an optimal solution. Now the sub problems are coupled by the global consensus vector \me{\mbfa{\zeta}} maintaining the complete stacked  interference profile of all users in the system as
\begin{IEEEeqnarray}{rCl}
\mbfa{\zeta} &=& \left [ \zeta_{1,\bar{\mc{U}}_{1}(1),1,1}, \dotsc, \zeta_{L,\bar{\mc{U}}_{1}(1),1,1}, \dotsc, \zeta_{L,\bar{\mc{U}}_{1}(|\bar{\mc{U}}_{1}|),1,1},  \right. \nonumber \\
&\quad& \left . \dotsc, \zeta_{L,\bar{\mc{U}}_{N_B}(|\bar{\mc{U}}_{N_B}|),1,N_B}, \dotsc, \zeta_{L,\bar{\mc{U}}_{N_B}(|\bar{\mc{U}}_{N_B}|),N,N_B} \right ] \IEEEyessubnumber \\
n_{b_k} &=& |\mbfa{\zeta}^{\set{b_k}}| = N L \sum_{b \in \mc{B}} \left | \bar{\mc{U}}_b\right | \IEEEyessubnumber
\end{IEEEeqnarray}

@@

\subsubsection*{Convergence}
The convergence of the \ac{ADMM} method follows the same argument as the centralized algorithm if each distributed algorithm is allowed to converge to the optimal value for a fixed \ac{SCA} point. Since subproblem solved at each \ac{BS} is convex, the \ac{ADMM} method converges to the optimal value \cite{boyd2011distributed} for a given \ac{SCA} point. The receive beamformers are updated at each \ac{SCA} update provides monotonic increase in the objective function, since the \ac{MMSE} receive beamformers are optimal for the fixed transmit precoders obtained by solving the subproblems until convergence. The algorithmic representation of the \ac{ADMM} based approach for decentralization is given in Algorithm \ref{algo-3}.
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{C}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=0} and the transmit precoders \me{\tilde{\mbf{m}}_{l,k,n}} randomly satisfying the total power constraint \eqref{eqn-4.3}}
 update \me{\mvec{w}{l,k,n}} with \eqref{eqn-10} and \me{\tilde{\mbf{u}}_{l,k,n}} with \eqref{eqn-8} using \me{\tilde{\mbf{m}}_{l,k,n}} \;
 initialize the global interference vectors \me{\mbfa{\zeta}^{(0)} = \mbfa{0}^{\mathrm{T}}} \;
 initialize the interference threshold \me{\mbfa{\nu}^{\set{b}(0)} \forall b \in \mc{B} = 0} \;
 for each \ac{BS} \me{b \in \mc{B}}, perform the following procedure \;
 \Repeat{convergence or \me{i \geq I_{\max}}}{
    initialize \me{j=0} \;
    \Repeat{convergence or \me{j \geq J_{\max}}}{
        solve for the transmit precoders \me{\mvec{M}{k,n}} and the local interference \me{\mbfa{\zeta}^{\set{b}}} using \eqref{eqn-dual-3} \;
        exchange \me{\mbfa{\zeta}^{\set{b}(j)}} across the coordinating \acp{BS} in \me{\mc{B}} \;
		update the dual variables in \me{\mbfa{\nu}^{\set{b}(j+1)}} using \eqref{dual-sg-update} \;
        update the global interference vector \me{\mbfa{\zeta}^{(j+1)}} using \eqref{zeta_update} \;
		\me{j = j + 1} \;
    }
    update the receive beamformers \me{\mvec{w}{l,k,n}} using \eqref{eqn-10} by exchanging the recent precoders \me{\mvec{m}{l,k,n}} \;
    exchange the receive precoders \me{\mbf{W}_{k,n}} \me{\forall k \in \mc{U}_b} among the \acp{BS} in \me{\mc{B}} \;
    update \me{\tilde{p}_{l,k,n}}, \me{\tilde{q}_{l,k,n}} and \me{\tilde{\beta}_{l,k,n}} with the recent precoders using \eqref{eqn-wsrm-expr} and \eqref{eqn-6.3} for the \ac{SCA} approach (or)\;
    update \me{\tilde{\epsilon}_{l,k,n}} with the recent precoders using \eqref{eqn-mse-2.3} with equality for the \ac{MSE} formulation approach \;
    $i = i + 1$ \;
  }
 \caption{Decentralization via \ac{ADMM} for \acs{JSFRA} scheme}
  \label{algo-3}
\end{algorithm}

>>>>

In this section, we discuss the \ac{ADMM} decomposition method, which is basically based on the \acl{DD}, but shows better convergence properties. In contrast to the \acl{PD} problem, the \ac{ADMM} method relaxes the interference constraints by including it in the objective function of each subproblem with a penalty pricing \cite{palomar2006tutorial,boyd2011distributed}. In order to decouple the problem \eqref{eqn-decent-1}, the coupling variables \me{\zeta_{l,k,n,b}} in \eqref{eqn-decent-3} are replaced by the respective local copies \me{\mbfa{\zeta}^{\set{b}}, \forall b \in \mc{B}}, which are then solved for an optimal solution. Now the sub problems are coupled by the global consensus vector \me{\mbfa{\zeta}} maintaining the complete stacked  interference profile of all users in the system as
\begin{IEEEeqnarray}{rCl}
\mbfa{\zeta} &=& \left [ \zeta_{1,\bar{\mc{U}}_{1}(1),1,1}, \dotsc, \zeta_{L,\bar{\mc{U}}_{1}(1),1,1}, \dotsc, \zeta_{L,\bar{\mc{U}}_{1}(|\bar{\mc{U}}_{1}|),1,1},  \right. \nonumber \\
&\quad& \left . \dotsc, \zeta_{L,\bar{\mc{U}}_{N_B}(|\bar{\mc{U}}_{N_B}|),1,N_B}, \dotsc, \zeta_{L,\bar{\mc{U}}_{N_B}(|\bar{\mc{U}}_{N_B}|),N,N_B} \right ] \IEEEyessubnumber \\
n_{b_k} &=& |\mbfa{\zeta}^{\set{b_k}}| = N L \sum_{b \in \mc{B}} \left | \bar{\mc{U}}_b\right | \IEEEyessubnumber
\end{IEEEeqnarray}

@@

\subsubsection*{Convergence}
The convergence of the \ac{ADMM} method follows the same argument as the centralized algorithm if each distributed algorithm is allowed to converge to the optimal value for a fixed \ac{SCA} point. Since subproblem solved at each \ac{BS} is convex, the \ac{ADMM} method converges to the optimal value \cite{boyd2011distributed} for a given \ac{SCA} point. The receive beamformers are updated at each \ac{SCA} update provides monotonic increase in the objective function, since the \ac{MMSE} receive beamformers are optimal for the fixed transmit precoders obtained by solving the subproblems until convergence. The algorithmic representation of the \ac{ADMM} based approach for decentralization is given in Algorithm \ref{algo-3}.
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=0} and the transmit precoders \me{\tilde{\mbf{m}}_{l,k,n}} randomly satisfying the total power constraint \eqref{eqn-4.3}}
 update \me{\mvec{w}{l,k,n}} with \eqref{eqn-10} and \me{\tilde{\mbf{u}}_{l,k,n}} with \eqref{eqn-8} using \me{\tilde{\mbf{m}}_{l,k,n}} \;
 initialize the global interference vectors \me{\mbfa{\zeta}^{(0)} = \mbfa{0}^{\mathrm{T}}} \;
 initialize the interference threshold \me{\mbfa{\nu}^{\set{b}(0)} \forall b \in \mc{B} = 0} \;
 for each \ac{BS} \me{b \in \mc{B}}, perform the following procedure \;
 \Repeat{convergence or \me{i \geq I_{\max}}}{
    initialize \me{j=0} \;
    \Repeat{convergence or \me{j \geq J_{\max}}}{
        solve for the transmit precoders \me{\mvec{M}{k,n}} and the local interference \me{\mbfa{\zeta}^{\set{b}}} using \eqref{eqn-dual-3} \;
        exchange \me{\mbfa{\zeta}^{\set{b}(j)}} across the coordinating \acp{BS} in \me{\mc{B}} \;
		update the dual variables in \me{\mbfa{\nu}^{\set{b}(j+1)}} using \eqref{dual-sg-update} \;
        update the global interference vector \me{\mbfa{\zeta}^{(j+1)}} using \eqref{zeta_update} \;
		\me{j = j + 1} \;
    }
    update the receive beamformers \me{\mvec{w}{l,k,n}} using \eqref{eqn-10} by exchanging the recent precoders \me{\mvec{m}{l,k,n}} \;
    exchange the receive precoders \me{\mbf{W}_{k,n}} \me{\forall k \in \mc{U}_b} among the \acp{BS} in \me{\mc{B}} \;
    update \me{\tilde{p}_{l,k,n}}, \me{\tilde{q}_{l,k,n}} and \me{\tilde{\beta}_{l,k,n}} with the recent precoders using \eqref{eqn-wsrm-expr} and \eqref{eqn-6.3} for the \ac{SCA} approach (or)\;
    update \me{\tilde{\epsilon}_{l,k,n}} with the recent precoders using \eqref{eqn-mse-2.3} with equality for the \ac{MSE} formulation approach \;
    $i = i + 1$ \;
  }
 \caption{Decentralization via \ac{ADMM} for \acs{JSFRA} scheme}
  \label{algo-3}
\end{algorithm}

### 2014-08-13 Report/ieee_journal.tex
completed_journal
Finalized before reviewing - 2
modified for review

\begin{abstract}
%\input{abstract}
\end{abstract}

\section{Introduction} \label{sec-1}
\input{introduction}

@@

\section*{Acknowledgment}
This work has been supported by the Finnish Funding Agency for Technology and Innovation (Tekes), Nokia Solutions Networks, Xilinx Ireland, Renesas Mobile Europe, Academy of Finland.

\section{Conclusions} \label{sec-6}

>>>>

\begin{abstract}
\input{abstract}
\end{abstract}

\acresetall
\section{Introduction} \label{sec-1}
\input{introduction}

@@

\section*{Acknowledgment}
This work has been supported by the Finnish Funding Agency for Technology and Innovation (Tekes), Nokia Solutions Networks, Xilinx Ireland, Academy of Finland.

\acresetall
\section{Conclusions} \label{sec-6}
\input{conclusions}

### 2014-08-13 Report/introduction.tex
modified for review

In a network with multiple \acp{BS} serving multiple users (\acs{MU}), the main driving factor for the transmission are the packets waiting at each \ac{BS} corresponding to the different users present in the network. These available packets are transmitted over the shared wireless resources subject to certain system limitations and constraints. In this work, we consider the problem of transmit design over the space-frequency resources provided by a \ac{MU-MIMO} \ac{OFDM} framework in the downlink broadcast transmission to minimize the number of queued packets. Since the space-frequency resources are shared by multiple users associated with different \acp{BS}, the problem of interest can be viewed as a resource allocation to minimize the total number of backlogged packets of all \acp{BS}.

@@

In this paper, we consider the problem of precoder design across the space-frequency resources to minimize the total number of queued packets of all \acp{BS}. For this highly nonconvex problem, we first propose two centralized methods. In the first method, we relax the nonconvex constraint by the first order taylor approximation around an operating point, which is updated in an iterative manner until convergence or to a certain accuracy. In the second method, we reformulate the \ac{JSFRA} problem using the \ac{MSE} equivalence with the rate expression to solve for the optimal precoders. For distributed implementation, we further proposed decentralized approaches based on primal and \ac{ADMM} scheme to identify the precoders independently across the \acp{BS} by exchanging limited information via back-haul. We also proposed an iterative algorithm by solving the \ac{KKT} equations, which can be implemented easily in a distributed manner.

>>>>

In a network with multiple \acp{BS} serving multiple users (\acs{MU}), the main driving factor for the transmission are the packets waiting at each \ac{BS} corresponding to the different users present in the network. These available packets are transmitted over the shared wireless resources subject to certain system limitations and constraints. In this work, we consider the problem of transmit design over the space-frequency resources provided by the \acs{MIMO}-\ac{OFDM} framework in the downlink broadcast transmission to minimize the number of queued packets. Since the space-frequency resources are shared by multiple users associated with different \acp{BS}, the problem of interest can be viewed as a resource allocation to minimize the total number of backlogged packets of all \acp{BS}.

@@

In this paper, we consider the problem of precoder design across the space-frequency resources to minimize the total number of queued packets of all \acp{BS}. For this highly nonconvex problem, we first propose two centralized methods. In the first method, we relax the nonconvex constraint by the first order Taylor approximation around an operating point, which is updated in an iterative manner until convergence or to a certain accuracy. In the second method, we reformulate the \ac{JSFRA} problem using the \ac{MSE} equivalence with the rate expression to solve for the optimal precoders. For distributed implementation, we further proposed decentralized approaches based on primal and \ac{ADMM} scheme to identify the precoders independently across the \acp{BS} by exchanging limited information via back-haul. We also proposed an iterative algorithm by solving the \ac{KKT} equations, which can be implemented easily in a distributed manner.

### 2014-08-13 Report/mse_formulation.tex
modified for review

In this section, we solve the \ac{JSFRA} problem by exploiting the equivalence between the \ac{MSE} and the achievable capacity for the receivers designed based on the \ac{MMSE} criterion \cite{viswanath1999optimal,mse_duality}. The \ac{MSE} for the data symbol, represented as \me{\epsilon_{l,k,n}}, is given by
\begin{subeqnarray} \label{mse-error}
\epsilon_{l,k,n} &=& \mathbb{E} \big [ ( d_{l,k,n} - \hat{d}_{l,k,n} )^2 \big ] \\
 &=&  \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 + \sum_{\mathclap{(j,i) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_i,i,n} \mvec{m}{j,i,n} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}\|^2
\end{subeqnarray}
where \me{\mvec{m}{l,k,n}, \mvec{w}{l,k,n}} denotes the transmit and the receive beamformer and \me{\hat{d}_{l,k,n}} is the received symbol as in \eqref{eqn-1}. Now, replacing the receive beamformer in \eqref{mse-error} with the \ac{MMSE} receiver shown in \eqref{eqn-10}, we obtain the following relation between the \ac{MSE} and the \ac{SINR} as
\begin{equation} \label{eqn-mse-eq}
\epsilon_{l,k,n} = \dfrac{1}{1 + \gamma_{l,k,n}}
\end{equation}
where \me{\gamma_{l,k,n}} is the received \ac{SINR} as in \eqref{eq:SINR}. Using the equivalence in \eqref{eqn-mse-eq}, the \ac{WSRM} objective can be reformulated as the \ac{WMMSE} objective to obtain the precoders for the \ac{MU-MIMO} scenario as discussed in \cite{christensen2008weighted,wmmse_shi}.

@@

\subsubsection*{Convergence}
The convergence of the \ac{MSE} reformulation approach follows the same arguments detailed in \cite{wmmse_shi,christensen2008weighted}. Since the transmit precoders are obtained by minimizing the weighted \ac{MSE} criterion, it converges to the stationary \ac{KKT} point at each iteration. The receiver updates are made based on the \ac{MMSE} criterion, it provides the optimal receive beamformers for the given transmit precoders. Alternating between these two precoder designs guarantees the monotonicity of the objective function used in the problem \eqref{eqn-mse-2}. The \ac{MSE} relaxation method converges to the same stationary \ac{KKT} point as of the original \ac{WSRM} problem by making the graidents equal.

>>>>

In this section, we solve the \ac{JSFRA} problem by exploiting the equivalence between the \ac{MSE} and the achievable capacity for the receivers designed based on the \ac{MMSE} criterion \cite{viswanath1999optimal,mse_duality}. The \ac{MSE} for the data symbol, represented as \me{\epsilon_{l,k,n}}, is given by
\begin{subeqnarray} \label{mse-error}
\epsilon_{l,k,n} &=& \mathbb{E} \big [ ( d_{l,k,n} - \hat{d}_{l,k,n} )^2 \big ] \\
 &=&  \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 + \sum_{\mathclap{(j,i) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_i,i,n} \mvec{m}{j,i,n} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}\|^2
\end{subeqnarray}
where \me{\mvec{m}{l,k,n}, \mvec{w}{l,k,n}} denotes the transmit and the receive beamformer and \me{\hat{d}_{l,k,n}} is the received symbol as in \eqref{eqn-1}. Now, replacing the receive beamformer in \eqref{mse-error} with the \ac{MMSE} receiver shown in \eqref{eqn-10}, we obtain the following relation between the \ac{MSE} and the \ac{SINR} as
\begin{equation} \label{eqn-mse-eq}
\epsilon_{l,k,n} = \dfrac{1}{1 + \gamma_{l,k,n}}
\end{equation}
where \me{\gamma_{l,k,n}} is the received \ac{SINR} as in \eqref{eq:SINR}. Using the equivalence in \eqref{eqn-mse-eq}, the \ac{WSRM} objective can be reformulated as the \ac{WMMSE} objective to obtain the precoders for the \acs{MU}-\acs{MIMO} scenario as discussed in \cite{christensen2008weighted,wmmse_shi}.

@@

\subsubsection*{Convergence}
The convergence of the \ac{MSE} reformulation approach follows the same arguments detailed in \cite{wmmse_shi,christensen2008weighted}. Since the transmit precoders are obtained by minimizing the weighted \ac{MSE} criterion, it converges to the stationary \ac{KKT} point at each iteration. The receiver updates are made based on the \ac{MMSE} criterion, it provides the optimal receive beamformers for the given transmit precoders. Alternating between these two precoder designs guarantees the monotonicity of the objective function used in the problem \eqref{eqn-mse-2}. The \ac{MSE} relaxation method converges to the same stationary \ac{KKT} point as of the original \ac{WSRM} problem by making the gradients equal.

### 2014-08-13 Report/problem_formulation.tex
Signed-off-by: Ganesh <gvenkatr@ee.oulu.fi>
modified for review

The problem defined in \eqref{q_gen_sum} ignores the second order term arising from the Lyapunov drift minimization objective by the limiting it to a constant value. Eq. \eqref{eqn-4.2} provides similar expression when the exponent is set to be \me{\ell_{q=2}} as
\begin{equation}\label{pf-1}
\underset{t_k}{\text{minimize}} \, \sum_k \, v_k^2 = \underset{t_k}{\text{minimize}} \, \sum_k \, Q_k^2 - 2 \, Q_k t_k + t_k^2.
\end{equation}
It is evident that \eqref{pf-1} is equivalent to \eqref{backpreassure-approx} if the second order terms are ignored. Limiting \me{t_k^2} by a constant value, the \ac{Q-WSRM} formulation requires the explicit rate constraint \eqref{eqn-3.1.4} to avoid the resource wastage in the form of over allocation. In the current queue deviation formulation, the explicit rate constraint is not needed, since it is handled by the objective function itself. In contrast to the \ac{WSRM} formulation, the \ac{JSFRA} and the \ac{Q-WSRM} problems include the sub-channels jointly to achive an efficient allocation by identifying the optimal space-frequency resource for each user in the system. The queue deviation objective provides an alternate and a more efficient way to perform the resource allocation as compared to the \ac{Q-WSRM} approach.

>>>>

The problem defined in \eqref{q_gen_sum} ignores the second order term arising from the Lyapunov drift minimization objective by the limiting it to a constant value. Eq. \eqref{eqn-4.2} provides similar expression when the exponent is set to be \me{\ell_{q=2}} as
\begin{equation}\label{pf-1}
\underset{t_k}{\text{minimize}} \, \sum_k \, v_k^2 = \underset{t_k}{\text{minimize}} \, \sum_k \, Q_k^2 - 2 \, Q_k t_k + t_k^2
\end{equation}
It is evident that \eqref{pf-1} is equivalent to \eqref{backpreassure-approx} if the second order terms are ignored. Limiting \me{t_k^2} by a constant value, the \ac{Q-WSRM} formulation requires the explicit rate constraint \eqref{eqn-3.1.4} to avoid the resource wastage in the form of over allocation. In the current queue deviation formulation, the explicit rate constraint is not needed, since it is handled by the objective function itself. In contrast to the \ac{WSRM} formulation, the \ac{JSFRA} and the \ac{Q-WSRM} problems include the sub-channels jointly to achieve an efficient allocation by identifying the optimal space-frequency resource for each user in the system. The queue deviation objective provides an alternate and a more efficient way to perform the resource allocation as compared to the \ac{Q-WSRM} approach.
